---
title: "epic-annotation"
author: "ks"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


Get Manifest from Illumina. Column names:
https://knowledge.illumina.com/microarray/general/microarray-general-reference_material-list/000001568

```{r read-annot}
idat_dir <- c("~kims/Google Drive/My Drive/shibata/Shibata EPIC Data/datajy")
mft <- read_csv(file.path(idat_dir,
                          "infinium-methylationepic-v-1-0-b5-manifest-file.csv"),
                          skip=7)
dim(mft)
```

Now subset to CpGs only

```{r subset-cgs}
table(substr(mft$Name,1,2))

mft  <- mft  %>%
          filter(substr(Name,1,2)=="cg") %>%
          arrange(Name)
```

For gene and region annotation, we're going to use just the first entry for each CpG.

```{r gene-name}
xlist <-  strsplit(mft$UCSC_RefGene_Name,";")
ucsc_refgene_name  <- map_chr(xlist,pluck,1)
length(ucsc_refgene_name)
ucsc_refgene_name[1:3]
```

```{r gene-group}
xlist <-  strsplit(mft$UCSC_RefGene_Group,";")
ucsc_refgene_group  <- map_chr(xlist,pluck,1)
ucsc_refgene_group[1:3]
```

```{r cpg-annot}
cpgannot <- cbind.data.frame(IlmnID = mft$IlmnID,
                             chr = mft$CHR,
                             mapinfo = mft$MAPINFO,
                             ucsc_refgene_name,
                             ucsc_refgene_group,
                             Relation_to_UCSC_CpG_Island =
                               mft$Relation_to_UCSC_CpG_Island,
                             Regulatory_Feature_Group =
                               mft$Regulatory_Feature_Group)
```


## nCpG per gene

Count by Group
Then reformat data to wide format.
```{r count-CpGs}
genesum <- cpgannot  %>%
      #select(CHR,MAPINFO,UCSC_RefGene_Name) %>%
      group_by(ucsc_refgene_name,ucsc_refgene_group)  %>%
      mutate( n = 1)   %>%
      summarise(nCpG = sum(n))
tail(genesum)
genesum <- genesum[!is.na(genesum$ucsc_refgene_name),]
tail(genesum)
```

Reformat data.

```{r}
ucscnf <- genesum %>% 
    pivot_wider(names_from = ucsc_refgene_group, 
                values_from = nCpG)
head(ucscnf)
```

Summarize distribution of the number of CpGs in a gene region.
```{r disns}
nc <- ncol(ucscnf)
for (i in 2:nc) {
  #print(colnames(ucscnf)[i])
  print(summary(ucscnf[,i],na.action=omit))
  }
```


Get totals for each gene (apply to dataset without gene name)

visualize the counts... how?   for ucsc refgene.  There are other annotations, but this is the definitions Darryl has been using in our papers.

```{r si}
sessionInfo()
```


