---
title: "Get paired DNAm differences"
author: "ks"
date: "2023-12-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sesame)
library(GenomicRanges)
library(tidyverse)
library(parallel)
```



## Process Idats

Tumor M Plate 1350   
Tumor MA,MB barcodes: 200705860031_R04C01,
200705860031_R05C01

Tumor I Plate 1688
Tumor IA1, IB2 barcodes:  204367410162_R03C01
204367410162_R04C01

Directory:
~kims/Google Drive/My Drive/shibata/Shibata EPIC Data/datajy.

```{r dir}
idat_dir <- c("~kims/Google Drive/My Drive/shibata/Shibata EPIC Data/datajy")
```


Read in the beta values and compute for each sample, the fraction of features with missing data.

```{r Sesame}
betas = openSesame(idat_dir, func = getBetas) # getBetas is the default
#nrow(betas)
betas <-  betas[substr(rownames(betas),1,2)=="cg" ,]
nrow(betas)
apply(is.na(betas),2,mean)
```

This is a lot of missing data.  It must be applying masks.  Let's filter on ucsc_refgene_name, and compute the proportion of features masked in each gene.
First we have to merge in the refseq gene names. Then save CpGs in genes.


```{r filter-to-refseqgenes}
load("data/cpgannot.rda")
betas <- cbind.data.frame(IlmnID = rownames(betas),
                          betas)
beta2 <- merge(betas, cpgannot) 
beta_gene <- beta2 %>%
        filter(!is.na(ucsc_refgene_name))
dim(beta_gene)
apply(is.na(beta_gene[,c("200705860031_R04C01","200705860031_R05C01",
                        "204367410162_R03C01","204367410162_R04C01")]),2,mean)
```
Sesame masks ~12\% of CGs in Refseq genes due to poor quality probe design. 


Compute differences, means and absolute differences.
```{r}
ec <- cbind.data.frame( IlmnID = beta_gene$IlmnID,
                            Md = beta_gene[,"200705860031_R04C01"]-
                                  beta_gene[,"200705860031_R05C01"],
                          Mpwd = abs(beta_gene[,"200705860031_R04C01"]-
                                  beta_gene[,"200705860031_R05C01"]),
                            Ma = (beta_gene[,"200705860031_R04C01"]+
                                  beta_gene[,"200705860031_R05C01"])/2,
                            Id =  beta_gene[,"204367410162_R03C01"]-
                                  beta_gene[,"204367410162_R04C01"],
                          Ipwd = abs(beta_gene[,"204367410162_R03C01"]-
                                  beta_gene[,"204367410162_R04C01"]),
                            Ia = (beta_gene[,"204367410162_R03C01"]+
                                   beta_gene[,"204367410162_R04C01"])/2)
head(ec)
```

```{r miss-dat}
nmis <- apply(is.na(ec[,2:5]),1,sum)
table(nmis)
```

Remove features that were NAs for all samples.
```{r filter}
ec <- ec[nmis!=4,]
save(ec,file="data/ec.rdat")
```


plot M vs A

```{r  ggplot, cache=TRUE}
ggplot(ec,aes(x=Ia,y=Id )) +
        stat_density2d(aes(fill = after_stat(density)^0.25), geom = "tile",
                       contour = FALSE, n=200, show.legend=FALSE) +
        scale_fill_continuous(low="white", high="dodgerblue4")
```


```{r  ggplot-pwd, cache=TRUE}
ggplot(ec,aes(x=Ia,y=Ipwd )) +
        stat_density2d(aes(fill = after_stat(density)^0.25), geom = "tile",
                       contour = FALSE, n=200, show.legend=FALSE) +
        scale_fill_continuous(low="white", high="dodgerblue4") +
       geom_smooth(color="red")
```

```{r sI}
sessionInfo()
```

