---
title: "Visualize Epi Conservation Data"
author: "ks"
date: "2023-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sesame)
library(GenomicRanges)
library(tidyverse)
library(parallel)
```

Load CpG level data on pwds.
```{r pwd}
load("data/ec.rdat")
load("data/crcdev.rda")
```
Run epic-annot.Rmd to get annotation file: cpgannot.

Combine with annotation
```{r merge-ec-cpgannot}
annoec <- merge(x=cpgannot, 
                y=ec,
                all.x = FALSE,
                all.y = TRUE) %>%
           arrange(chr,mapinfo)
```

Let's explore genes with no expression & require 4 CpGs.
```{r}
nodev <- crcdev %>%
     filter(jdev < 0.5) %>%
     arrange(name)
nodev
```



```{r count-CpGs}
cpgcount <- cpgannot  %>%
      group_by(ucsc_refgene_name)  %>%
      mutate(n = 1)   %>%
      summarise(nCpG = sum(n))
tail(cpgcount)

cpgcount <- cpgcount %>%
      filter(nCpG > 3)

cpgcount <- cpgcount[!is.na(cpgcount$ucsc_refgene_name),]
tail(cpgcount)

cpgcount
```

```{r noexpr-}
subset <- merge(x = nodev,
                y = cpgcount,
                by.x = "name",
                by.y = "ucsc_refgene_name")
subset
```



Write function to get all CpGs for a single gene and plot beta value vs ranked location. Individually explored: ABCC11, AADACL2-AS1 (binning does nothing)
```{r}
annoec %>% 
    filter(ucsc_refgene_name=="ADARB2") %>%
    ggplot(aes(x=mapinfo,y=Mpwd)) +
      geom_point()
```

```{r}
annoec %>% 
    filter(ucsc_refgene_name=="ADARB2") %>%
    summarize(aMpwd = mean(Mpwd,na.rm=T), 
              sMpwd = sd(Mpwd,na.rm=T),
              aIpwd = mean(Ipwd,na.rm=T),
              sIpwd = sd(Ipwd,na.rm=T))
```
Smoothing
```{r}
dat <- annoec %>% 
    filter(ucsc_refgene_name=="ADARB2") 
max(dat$mapinfo)-min(dat$mapinfo)
```
```{r}
gc <- boundedClusterMaker(dat$chr, pos=dat$mapinfo, assumeSorted = FALSE,
maxClusterWidth = 1500, maxGap = 500)
gc
```

Average by cluster, then gene.
```{r}
add_column(dat,gc) %>%
    group_by(gc) %>%
    summarize(ablMpwd = mean(Mpwd,na.rm=T)) %>%
    summarize(Mec = mean(ablMpwd,na.rm=T), sdM =sd(ablMpwd,na.rm=T))
```



```{r}
annoec %>% 
    filter(ucsc_refgene_name=="ADARB2") %>%
    summary(Mpwd)
```

For ABCC11 (29 CpGs) it looks like 75th percentile would have been better than the mean. Also better than averaging over bin averages (16 bins).

This doesn't improve slope of Dev vs. PWD summary. What does improve it is requiring more CpGs than just 4. Go back and get the deviance without the jitter (jdev). 


```{r sI}
sessionInfo()
```

